---
title: "FacileAnalysis: genes to pathways using the edgeR quasi-likelihood pipeline"
author: "Steve Lianoglou"
date: "March 4, 2020"
output:
  rmarkdown::html_document:
    css: "custom.css"
    highlight: pygments
    toc: true
    toc_float: true
    fig_width: 6
    df_print: paged
bibliography: library.bib
vignette: >
  %\VignetteIndexEntry{FacileAnalysis: genes to pathways using the edgeR quasi-likelihood pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  # code or die
  echo = TRUE,
  # minimize verbosity
  warning = FALSE, message = FALSE,
  # dpi = 150, # for hires images
  comment = "#>")
set.seed(0xFEED)
options(digits = 3)
```

## Overview

This vignette provides 1:1 comparisons of the results generated using this
toolkit vs the standard edgeR quasi likelihood maneuvers to show that they
are equivalent.

It's only meant to show equivalence of results, not as a tutorial of the
FacileAnalysis package: [reference the sister vignette][facileqlf] for that.

[facileqlf]: FacileRnaSeqGeneEdgeRQL.Rmd

```{r init, warning=FALSE, message=FALSE}
library(dplyr)
library(FacileData)
library(FacileBiocData)
library(FacileAnalysis)
library(edgeR)
library(ggplot2)
theme_set(theme_bw())
```

## Assemble the DGEList

```{r}
source(system.file("scripts", "edgeRQLF-workflow.R", package = "FacileAnalysis"))
y <- assemble_edgeRQLF_DGEList()
y <- calcNormFactors(y)

# Do gene filtering

```


## Reduced Dim

```{r}
pch <- c(0,1,2,15,16,17)
colors <- rep(c("darkgreen", "red", "blue"), 2)
plotMDS(y, col = colors[y$samples$group], pch=pch[y$samples$group])
legend("topleft", legend = levels(y$samples$group), pch = pch, 
       col = colors, ncol = 2)
```

Curently, the only dimmensionality reduction we support is PCA, so we'll plot
that with a slightly different aesthetic mapping scheme.

```{r}
pca <- fpca(f, ntop = 500)
status.color <- c(lactating = "darkgreen", pregnant = "red", virgin = "blue")
viz(pca, color_aes = "Status", shape_aes = "CellType",
    color_map = status.color, shape_map = colors$group, width = 600)
```

## Differential Expression Analysis

```{r}
y <- estimateDisp(y, design, robust=TRUE)
```

```{r}
plotBCV(y)
```


```{r}
fit <- glmQLFit(y, design, robust=TRUE)
```

```{r}
B.LvsP <- makeContrasts(B.lactating-B.pregnant, levels=design)
res <- glmQLFTest(fit, contrast=B.LvsP)
```

```{r}
qlf.tt <- as.data.frame(topTags(res, n = Inf))
```

```{r}
facile.res <- y %>% 
  flm_def("group", numer = "B.lactating", denom = "B.pregnant") %>% 
  fdge(method = "edgeR-qlf", filter = FALSE)
facile.tt <- tidy(facile.res)
```

```{r}
cmp <- full_join(
  dplyr::select(qlf.tt, feature_id, symbol, logFC:FDR),
  dplyr::select(facile.tt, feature_id, logFC:padj),
  by = "feature_id", suffix = c(".edger", ".facile"))
```

```{r}
ggplot(cmp, aes(x = -log10(PValue), y = -log10(pval))) +
  geom_point() +
  geom_abline(slope = 1, yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "p-value comparison between edgeR and facile::fdge('edgeR-qlf')",
    x = "-log10(edgeR pvalue)",
    y = "-log10(facile QLF pvalue)")
```

```{r}
is.de <- decideTestsDGE(res)
plotMD(res, status=is.de, values=c(1,-1), col=c("red","blue"),
       legend="topright")
```

The MA plots in facile are interactive, so we don't want to overload the
web-browser. Here we'll chose to show a sample of both significant and
"background" genes.

```{r}
bg.genes <- filter(facile.tt, padj > 0.1) %>% sample_n(200)
fg.genes <- filter(facile.tt, padj < 0.1) %>% sample_n(1000)
viz(facile.res, type = "MA", features = bg.genes, highlight = fg.genes)
```

```{r}
viz(facile.res, type = "volcano", features = bg.genes, 
    highlight = head(fg.genes, 100))
```

### Differential expression above a fold-change threshold

```{r}
tr <- glmTreat(fit, contrast=B.LvsP, lfc=log2(1.5))
tr.tt <- as.data.frame(topTags(tr, n = Inf))
```

We'll re-use the model from the original analsis and tweak the fdge parameters.

```{r}
facile.thresh <- facile.res %>% 
  model() %>% 
  fdge(method = "edgeR-qlf", treat_lfc = log2(1.5), filter = FALSE)
facile.tt.thresh <- tidy(facile.thresh)
```

```{r}
cmp.thresh <- full_join(
  dplyr::select(tr.tt, feature_id, symbol, logFC:FDR),
  dplyr::select(facile.tt.thresh, feature_id, logFC:padj),
  by = "feature_id", suffix = c(".edger", ".facile"))
```

```{r}
ggplot(cmp.thresh, aes(x = -log10(PValue), y = -log10(pval))) +
  geom_point() +
  geom_abline(slope = 1, yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "p-value comparison against threshold",
    x = "-log10(edgeR pvalue)",
    y = "-log10(facile QLF pvalue)")
```

:::note
The workflow goes into a heatmap section, but we will save that for later.
There are more interesting differential expression examples in the workflow
that we will continue here.
:::

### Analysis of Deviance

### Complicated Contrasts

Suppose we want to compare the three groups in the luminal population, i.e., virgin, pregnant and lactating. An appropriate contrast matrix can be created as shown below, to make pairwise comparisons between all three groups:

```{r edgeR-aova}
con <- makeContrasts(
  L.PvsL = L.pregnant - L.lactating,
  L.VvsL = L.virgin - L.lactating,
  L.VvsP = L.virgin - L.pregnant, levels = design)

anova.res <- glmQLFTest(fit, contrast = con) %>% 
  topTags(n = Inf) %>% 
  as.data.frame()
```


```{r}

a2 <- glmQLFTest(fit, coef = )
```

The analgous facile analysis would subset down to the `L` samples and run
an ANOVA across the `"Status"` covariate. This won't be exactly the same since 
the number of samples used for the Bayes-like step in the analysis isn't the
same, but should be close.

```{r}
anova.facile <- y %>% 
  filter_samples(CellType == "L") %>% 
  flm_def("Status") %>% 
  fdge(method = "edgeR-qlf", filter = FALSE)
```

```{r}
cmp.anova <- anova.res %>% 
  full_join(select(tidy(anova.facile), feature_id, F, pval, padj),
            by = "feature_id", suffix = c(".edger", ".facile"))
```

```{r}
ggplot(cmp.anova, aes(x = -log10(PValue), y = -log10(pval))) +
  geom_point() +
  geom_abline(slope = 1, yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "p-value comparison for ANOVA",
    x = "-log10(edgeR pvalue)",
    y = "-log10(facile QLF pvalue)")
```

## Complicated Contrasts

Basically testing an interaction term

```{r}
con <- makeContrasts(
  (L.lactating-L.pregnant)-(B.lactating-B.pregnant), 
  levels=design)
interaction.res <- glmQLFTest(fit, contrast=con) %>% 
  topTags(n = Inf) %>% 
  as.data.frame()
```

The facile way is to perform the two individual and simpler test separately
and then compare them.

We already have the `B.lactating - B.pregnant` results stored in `facile.res`.
We'll make another one for `L.lactating - L.pregnant`, and compare them.

```{r}
facile.L <- y %>% 
  flm_def("group", numer = "L.lactating", denom = "L.pregnant") %>% 
  fdge(method = "edgeR-qlf", filter = FALSE)
```

```{r}
i.facile <- compare(facile.L, facile.res)
```

## Heatmaps

Facile heatmaps are not supported yet, but we can show how to use iheatmapr
produce the same heatmap.

## Gene Set 

