---
title: "Facile (Gene) Set Expression Analysis"
author: "Steve Lianoglou"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to FacileAnalysis}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  # code or die
  echo = TRUE,
  # minimize verbosity
  warning = FALSE, message = FALSE,
  # dpi = 150, # for hires images
  comment = "#>")
set.seed(0xFEED)
```

**Under Construction**

The facile feature set enrichment analysis leans on the sparrow framework
for running GSEA methods, as well as providing gene set collections via 
`sparrow::GeneSetDb` objects.

This vignette assumes that you've got a `GeneSetDb` object in hand named `gdb`.

```{r, eval = FALSE}
library(FacileData)
library(FacileAnalysis)
samples <- exampleFacileDataSet() %>% 
  filter_samples(indication == "BLCA")
```

For now, I imagine (gene) set enerichment analyses will be performed over
**the result** of other analyses. This means that we can perform (G)SEA on:

1. The result of a differential expression analysis (`fdge`) using:
    - Pre-ranked methods like cameraPR and fgsea
    - Enrichment methods (goseq, hypergeometric testing)
   
   or some combination of the two.
   
    ```{r, eval=FALSE}
    dge.gsea <- samples %>% 
      flm_def(covariate = "sample_type", 
                 numer = "tumor", denom = "normal",
                 batch = "sex") %>% 
      fdge() %>% 
      ffsea(gdb, pc = 1, method = c("cameraPR", "goseq", rank_by = "t"))
    ```

2. The "highly loaded" genes from the first prinicpal component of a
   principal components analysis (`fpca`). Deafult GSEA methods are set for
   each analysis result type, too, if you don't want to specify them.

    ```{r, eval=FALSE}
    pc1.gsea <- samples %>% 
      fpca() %>% 
      ffsea(gdb, pc = 1)
    ```

In the immediate term, the GSEA functionality will lean heavily on the sparrow
package, including the bits that enable interactivity over the results
(ie. the sparrow.shiny package).

**Comparing (Gene) Set Enrichment Results**

To **compare** GSEA results, Thomas' idea of overlaying enrichment maps from
the two results on top of each other in some dynamic way may be the first
thing we try.




